---
title             : "Descriptive Statistics of the NRW80+ Dataset"
shorttitle        : "Descriptive Statistics of the NRW80+ Dataset"

author: 
  - name          : "Prof. Dr. Stephan Huber"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Im Mediapark 4e"
    email         : "stephan.huber@hs-fresenius.de"
    # role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
    #   - "Conceptualization"
    #   - "Writing - Original Draft Preparation"
    #   - "Writing - Review & Editing"
  # - name          : "Ernst-August Doelle"
  #   affiliation   : "1,2"
  #   role:
  #     - "Writing - Review & Editing"
  #     - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Fresenius University of Applied Science"
  - id            : "2"
    institution   : "Charlotte Fresenius University"

authornote: |
  All files related to that paper are hostes on github. see: [https://github.com/hubchev/ewa](https://github.com/hubchev/ewa). 

abstract: |
  In this paper, I illustrate the process of importing NRW80+ data [see @Zank2022Quality] into R. Additionally, I present descriptive statistics and graphical visualizations to gain insights into Likert-scaled surveys. The paper adheres to the APA style, implementing the R template provided by the 'papaja' package [@R-papaja].

keywords          : "papaja, NRW80+, descriptive statistics"

bibliography      : "references_NRW.bib"

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no
numbersections    : yes

figurelist        : yes
tablelist         : yes
footnotelist      : no

classoption       : "doc"
output            : papaja::apa6_pdf 
toc               : yes
---


```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
# rm(list = ls())
```


# Technical Note

In the following, I load (and install) packages that I use later on and I show information about my R session with `sessionInfo()`.

```{r, echo=TRUE, eval=TRUE}
# (Install and) load pacman package 
if (!require(pacman)) install.packages("pacman")

# load packages that are already installed and install packages that are not 
# installed yet and then load them:
pacman::p_load(tinylabels, 
               papaja,
               haven, 
               labelled, 
               janitor,
               skimr, 
               rstatix, 
               HH, 
               likert, 
               expss,
               tidyr, 
               ggstats,
               psych,
               sjlabelled,
               sjmisc,
               tidyverse, 
               MASS,
               dplyr,
               magick)

sessionInfo()
```

# Import Data

I host a R script on my GitHub account (see <https://raw.githubusercontent.com/hubchev/courses/main/scr/readin_GESIS.R>) that explains how to import the NRW80+ data. I have manually saved the data, `gesis.RData`, in a subfolder named `data`.

# How to Use the NRW80+ Data

## Load and Subset Data {#sec-load}

I load the data and select some variables that are of particular interest to me.

```{r echo=TRUE}
getwd()
load("../data/gesis.RData")
df <- dfdta |>
  select(starts_with("alter"), 
         ALT_agegroup, 
         ALT_sex, 
         famst1, famst7, 
         demtectcorr, 
         kogstat, 
         final, 
         geschlecht)

# Remove the common prefix from all variables
df <- df |> 
  mutate_all(~ set_label(., gsub("^Alternserleben: ", "", get_label(.))))
```

For simplification, let us focus on the questions that refer to the "Experience of Ageing" and create a new dataset `df_alterl` that contains only those questions:

```{r, echo=TRUE}
df_alterl <- df |> 
  select(alterl1, 
         alterl2, 
         alterl3, 
         alterl4, 
         alterl5, 
         alterl6, 
         alterl7, 
         alterl8, 
         alterl9, 
         alterl10) |> 
  drop_unused_labels() 

# to remove unused labels you can use drop_unused_labels():
df_alterl_un <- df_alterl |>
  drop_unused_labels()



summary(df_alterl)
```

## Get an Overview by Counting

### `table()` of R base

With the `table()` function, you can count how many observations of each unique value a variable contains:

```{r, echo=TRUE}
table(df_alterl$alterl1)
```

To do that for each variable of a dataset is easy using `~`, the pipe operator, and `map()` of the package `purrr` [@Wickham2023purrr]:

```{r, echo=TRUE}
df_alterl |> 
  map(~ table(.))
```

Using `proportions()` returns the conditional proportions:

```{r, echo=TRUE}
df_alterl |> 
  map(~ proportions(table(.)))
```

### `tabyl()` of `janitor`

With `tabyl()` which is part of `janitor` [@Firke2023janitor], we can get both nicely:

```{r, echo=TRUE}
df_alterl |> 
  tabyl(alterl1) 

df_alterl |> 
  map(~ tabyl(.))
```

### `frq()` of `sjmisc`

As the variables `df_alterl1` are factors. Thus, we can use the `sjmisc` package, see @Luedecke2018sjmisc and the cheatsheet of `sjmisc` <http://strengejacke.de/sjmisc-cheatsheet.pdf>. Also worth a reading is `browseVignettes("sjmisc")`.

For example, we can use `frq()` for nice frequency tables:

```{r, echo=TRUE}
df_alterl |> 
  map(~ frq(. , show.na = T))

```

## First Summary Statistics

###  Using `summary()` and `get_summary_stats()`

First, I am interested in the class of the data and some very basic summary statistics. 

```{r, echo=TRUE}
summary(df)

sumstat_alter <- df |> 
  get_summary_stats(
    alterl1, 
    alterl2, 
    alterl3, 
    alterl4, 
    alterl5, 
    alterl6, 
    alterl7, 
    alterl8, 
    alterl9, 
    alterl10,  
    type = "five_number")  

sumstat_alter
```

###  Using `psych::describe()`

A powerful alternative for descriptive summary statistics is provided by the function `describe()` of the `psych` package [@WilliamRevelle2023psych].

```{r, echo=TRUE}
sumstat_alter_psych <- df |>
  select(starts_with("alterl")) |> 
  psych::describe() |> 
  as_tibble(rownames="Question")  |> 
  select(-skew, -kurtosis, -range, -vars) 

sumstat_alter_psych
```

###  Using `summarize()` and the `tidyverse`

As you may be aware, the `tidyverse` package provides powerful and flexible functions such as `filter`, `select`, `group_by`, and `summarize`. Here is an example demonstrating how these functions can be utilized to create descriptive statistic tables:

```{r echo=TRUE}
descriptives <- dfdta |>  
  # filter(alterl1 > 0) |> 
  group_by(geschlecht)  |> 
  summarize(
    Mean = mean(alterl1)
    , Count = n()
    , SD = sd(alterl1)
    , Min = min(alterl1)
    , Max = max(alterl1)
  )

descriptives
```

## Make APA Tables using `apa_table()`

The R output shown above might not meet publishable standards as it requires proper formatting, including a table with a caption and adherence to APA rules. To achieve this, the `apa_table()` function is recommended, and further details can be found in @Aust2020papaja[section 4.2].


```{r tabrstatix, echo=TRUE}
apa_table(
  sumstat_alter
  , caption = "Summary Statistics: Experience of Ageing."
  , note = "This table contains all variables of `alterl*`."
  , escape = TRUE
  )
```

```{r tabsumstatalterpsych, echo = TRUE }
apa_table(
  sumstat_alter_psych
  , caption = "Summary Statistics: Experience of Ageing (psych)"
  , note = "This table contains all variables of `alterl*`."
  , escape = TRUE
)
```

```{r tabdescriptives, echo=TRUE}
apa_table(
  descriptives
  , caption = "Experience of Ageing: Valuing Relationships and Other People 
  More (By Gender)"
  , escape = TRUE
)
```

Table \@ref(tab:tabrstatix) was created with the function `get_summary_stats()` of the `rstatix` package [@Kassambara2023rstatix], Tables \@ref(tab:tabsumstatalterpsych) and \@ref(tab:tabsumstatalterpsychbal) were created with the function `describe()` of the `psych` package [@WilliamRevelle2023psych], and Table \@ref(tab:tabdescriptives) was created with the function `summarize()` of the `dplyr` package [@Wickham2023dplyr].



## Use the Likert Scale using `gglikert()`

We have seen that the data contain not only the five different (Likert scaled) answers. Thus, let us remove all values that have, in one or multiple questions, no answer of the Likert scale. The cleaned dataset is named `df_alterl_balance`.

```{r, echo=TRUE}
df_alterl_balance <- df_alterl %>%
  rowwise() %>%
  mutate(has_negative = ifelse(any(c(across(alterl1:alterl10)) < 0), 1, 0)) |> 
  filter(has_negative == 0) |> 
  select(starts_with("alter")) |> 
  as_tibble()
```

Using the `gglikert()` of the `ggstats` package [@Larmarange2023ggstats] allows us to draw nice graphs. I highly recommend reading the vignette of the package in the R documentation which you get with `vignette("gglikert")`.

Figures \@ref(fig:likertalterl1) and \@ref(fig:likertalterl3) shows the proportions of answers using `df_alterl` data and Figures \@ref(fig:likertalterl2) and \@ref(fig:likertalterl4) does so using the `df_alterl_balance` data whereby the latter to show the proportions stacked. Do you see any difference and can you explain the differences?

```{r likertalterl1,  fig.cap="Experience of Ageing: Proportions of Answers (df_alterl)", echo=TRUE}
gglikert(df_alterl, 
         exclude_fill_values = c("Weiß nicht", "Verweigert"),
         sort = "ascending"
         )
```

```{r likertalterl2,  fig.cap="Experience of Ageing: Proportions of Answers (df_alterl_balance)", echo=TRUE}
gglikert(df_alterl_balance,
         sort = "ascending"
         )
```

```{r likertalterl3,  fig.cap="Experience of Ageing: Proportions of Answers - Stacked (df_alter)", echo=TRUE}
gglikert_stacked(df_alterl,
                 sort = "ascending",
                 sort_method = "mean"
                 )
```

```{r likertalterl4,  fig.cap="Experience of Ageing: Proportions of Answers - Stacked (df_alterl_balance)", echo=TRUE}
gglikert_stacked(df_alterl_balance,
                 sort = "ascending",
                 sort_method = "mean"
                 )
```

As we are interested in the differences of the two samples, it makes sense to look as the summary statistics for the `df_alter_balance` sample. This is shown in Table \@ref(tab:tabsumstatalterpsychbal).

```{r tabsumstatalterpsychbal, echo=TRUE}
sumstat_alter_psych_bal <- df_alterl_balance |>
  psych::describe() |> 
  as_tibble(rownames="Question")  |> 
  dplyr::select(-skew, -kurtosis, -range, -vars) 

apa_table(
  sumstat_alter_psych_bal
  , caption = "Summary Statistics: Experience of Ageing - balanced (psych)"
  , note = "This table contains all variables of `alterl*` and only observations where all questions had been answered."
  , escape = TRUE
)
```

\clearpage

# Cross-Referencing in R Markdown

In adherence to the APA style guidelines [@Association2022Publication], it is imperative to reference all figures and tables by their respective numbers within the text. Avoid using generic phrases like "the table above" or "the figure below." Additionally, refrain from hard-coding the numbers for a more dynamic and standardized approach. @Xie2023R explains concisely how to do that with R Markdown, see: [https://bookdown.org/yihui/rmarkdown-cookbook/cross-ref.html](https://bookdown.org/yihui/rmarkdown-cookbook/cross-ref.html).

For example, I can refer to Table \@ref(tab:tabrstatix) with `\@ref(tab:tabrstatix)` because I have specified the corresponding label in the R code-chunk, see:

````{r, echo = TRUE, eval = FALSE} 

```{r tabrstatix, echo=TRUE}
apa_table(
  sumstat_alter
  , caption = "Summary Statistics: Experience of Ageing."
  , note = "This table contains all variables of `alterl*`."
  , escape = TRUE
  )
```

````

```

\clearpage

# Exercises

1.  With `knitr::purl("desc_NRW80.Rmd")` you can extract the whole R code from the R Markdown file and write it into the R script `desc_NRW80.R`. Try it.

2.  The dataset `gesis.RData` comes with two different tibbles: `dfsav` and `dfdta`. Is there a difference between these two when it comes to the statistics that are shown in this paper? To check that, rename the pdf file `desc_NRW80.pdf`, change the code in Section \@ref(sec-load) so that you are using the other data (`df <- dfdta |> ...` vs. `df <- dfsav |> ...`), knit the Rmd again, and compare the stats.

3.  Check possible differences in the `gglikert` plots when using `df_alterl_un` instead of `df_alterl`.

4.  The stats above show that dealing with missing or non-standard answers is a crucial thing. Please read chapter *Missing Values* of @Wickham2023R, see: <https://r4ds.hadley.nz/missing-values>.

5. The labels of the variables `alterl1:alterl10` have "Alternserleben: " at the beginning. This is not necessary and overloads the graphs. Please change the labels for all graphs using the following code in the respective place in the rmd and then knit it again.


```{r, echo=T, eval=F}
# Remove the common prefix from all variables
df <- df |> 
  mutate_all(~ set_label(., gsub("^Alternserleben: ", "", get_label(.))))
```



\clearpage

# References





