---
title: "Explain Weighted Mean"
author: "Stephan Huber"
date: "2023-12-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




The formula for the weighted mean is:
\[
\bar{x} = \frac{\sum_{i=1}^{n} w_i \cdot x_i}{\sum_{i=1}^{n} w_i}
\]

In this formula:

- $\bar{x}$ represents the weighted mean.
- $n$ is the number of observations.
- $w_i$ represents the weight for the $i$-th observation.
- $x_i$ represents the $i$-th observation value.


```{r echo=TRUE}
rm(list = ls())
wt <- c(5, 2, 2, 1)
x <- c(1, 2, 3, 4)
x_mean <- mean(x)
x_mean

x_wt_mean_1 <- weighted.mean(x, wt)
x_wt_mean_1
```


Let us calculate the weighted mean manually:

```{r echo=TRUE}
product <- wt*x
# Nominator
nom <- sum(product)
nom
# Denominator
denom <- sum(wt)
denom

x_wt_mean_2 <- nom/denom
x_wt_mean_2
```
## Exercise 1

Below you see an alternative way to calculate the weighted mean. Can you explain it?

```{r echo=TRUE}
w_div_sumw <- wt/denom
w_div_sumw
multi_ww_x <- w_div_sumw * x
multi_ww_x
x_wt_mean_3 <- sum(multi_ww_x)
x_wt_mean_3
```

## Exercise 2

a) Calculate mean, variance, weighted mean, and the variance of the weighted mean for `x`.

```{r echo=TRUE}
results <- data.frame(
  Statistic = c("Mean", "Variance", "Weighted Mean", "Weighted Variance"),
  Value = c(mean(x), var(x), weighted.mean(x, wt), sum(wt * (x - weighted.mean(x, wt))^2) / sum(wt))
)
print(results)
```

b) Do it again but use `tidyverse` and the function `summarize`.

```{r, message=FALSE}
if (!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse)
```
 

```{r, echo=TRUE}
df <- tibble(wt = wt, x = x)

summary_stats <- df %>%
  summarize(
    Mean = mean(x),
    Variance = var(x),
    Weighted_Mean = weighted.mean(x, wt),
    Weighted_Variance = sum(wt * (x - weighted.mean(x, wt))^2) / sum(wt)
  )

# Display the table
print(summary_stats)
``` 

## Exercise 3

```{r}
ModKogDat <- read.csv("../data/ModKogDat.csv", header=TRUE, sep=",") 
# Read in data
df_modkog <- ModKogDat |> 
  mutate(
    modus = as.factor(modus),
    kognition = as.factor(kognition)
    ) |> 
  group_by(modus, kognition) |> 
  mutate(
    id_num = cur_group_id(),
    m_str = substr(modus, 1, 2),
    k_str = substr(kognition, 1, 2),
    id = paste(m_str, k_str, sep = "_")
  ) |> 
  select(-m_str, -k_str) |> 
  tibble() |> 
  ungroup() 

pacman::p_load(stats)
interaction.plot(df_modkog$kognition, df_modkog$modus, response = df_modkog$dauer)
interaction.plot(df_modkog$modus, df_modkog$kognition, response = df_modkog$dauer)

anova2_model <- aov(dauer ~ modus + kognition + modus:kognition, data = df_modkog)
summary(anova2_model)

```

